---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(include = FALSE)

library(pacman)
p_load(lmerTest, tidyverse, simr)

big_model <- readRDS("big_model.rds")
test <- read_csv("test_clean.csv")
train <- read_csv("language_dev_asd_clean_2.csv")
```


## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
- Report the power analysis and comment on what you can (or cannot) use its estimates for.


```{r}
# "R is an autist" - Malte 2018

#old model 
TypesModel <- lmer(CHI_MLU ~ VISIT + I(VISIT^2) + verbalIQ_1 + MOT_MLU + tokens_MOT + 
                    (1+VISIT + I(VISIT^2)|SUBJ), data = train, REML = FALSE)

#running power analysis on diff. effects 
powerV = powerSim(TypesModel, simr::fixed("VISIT", method = "t"), nsim=20) #keep getting error. 
powerV #98-100, effect size = 0.47 

power2 = powerSim(TypesModel, simr::fixed("verbalIQ_1", method = "t"), nsim=20) #t-test. 
power2 #98-100, effect size = 0.061

power3 = powerSim(TypesModel, simr::fixed("MOT_MLU", method = "t"), nsim=20)
power3 #98-100, effect size = 0.35

power4 = powerSim(TypesModel, simr::fixed("tokens_MOT", method = "t"), nsim=20)
power4 #82.5 (76.51 - 87.50), effect size = -0.0002


```

with other model 

```{r}
#new model - basic - but we can expand..
#as factor diagnosis
train$Diagnosis <- as.factor(train$Diagnosis)
train$SUBJ <- as.factor(train$SUBJ)

HypesModel <- lmer(CHI_MLU ~ Diagnosis * VISIT + I(VISIT^2) +
                    (1+VISIT + (VISIT^2)|SUBJ), data = train, REML = FALSE)
HypesModel

#doing only 20 just for ease 
set.seed(911)
powerHype <- powerSim(HypesModel, simr::fixed("DiagnosisTD", method = "t"), nsim = 20)
powerHype #20% -0.21

powerVisit <- powerSim(HypesModel, simr::fixed("VISIT", method = "t"), nsim = 20)
powerVisit #100% - effect size = 0.41
```


### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- [GitHub] if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}
HypesModel
fixef(HypesModel)["VISIT"] <- 0.1 #what is the smallest beta coefficient that would be interesting
fixef(HypesModel)["DiagnosisTD"] <- -0.1 #what is the smallest beta coefficient that would be int.

HypesModel

#this does not work - however, it works for the complex model, so mayby I did something strange. 
library(simr)
powerCurveV = powerCurve(HypesModel,
fixed("VISIT"),along="SUBJ",nsim = 10)

powerCurveV = powerCurve(HypesModel,
fixed("DiagnosisTD"),along="SUBJ", nsim = 10)


powerCurveV 
plot(powerCurveV)

#simulating more participants 
#I think we should use makeLmer()

#other model
TypesModel
fixef(TypesModel)["VISIT"] <- 0.1 #what is the smallest beta coefficient that would be interesting
fixef(TypesModel)["MOT_MLU"] <- 0.1 #what is the smallest beta coefficient that would be
TypesModel

powerCurveV = powerCurve(TypesModel,
fixed("VISIT"),along="SUBJ", nsim = 10)

powerCurveV
plot(powerCurveV)

```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why
