---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(include = FALSE)

library(pacman)
p_load(lmerTest, tidyverse, simr)

big_model <- readRDS("big_model.rds")
test <- read_csv("test_clean.csv")
train <- read_csv("language_dev_asd_clean_2.csv")
```


## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

with other model 

```{r}
#new model - basic - but we can expand..
#as factor diagnosis
train$Diagnosis <- as.factor(train$Diagnosis)
train$SUBJ <- as.factor(train$SUBJ)

ggplot(train, aes(VISIT, CHI_MLU, color = Diagnosis))+
  geom_point()+
  geom_smooth()

#first model
HypesModel <- lmer(CHI_MLU ~ Diagnosis * VISIT + Diagnosis * I(VISIT^2) +
                    (1+VISIT +(VISIT^2)|SUBJ) , data = train, REML = FALSE)

#interaction effect?
inter_diag_visit <- powerSim(HypesModel, simr::fixed("DiagnosisTD:VISIT", method = "t"), nsim = 20)
inter_diag_visit #40% - effect size = -0.21

#second model 
PypesModel <- lmer(CHI_MLU ~ Diagnosis + VISIT + I(VISIT^2) + (1+VISIT +(VISIT^2)|SUBJ),
                   data = train, REML = FALSE)

fix_diag <- powerSim(PypesModel, simr::fixed("Diagnosis", method = "t"), nsim = 20)
fix_diag

fix_visit <- powerSim(PypesModel, simr::fixed("VISIT", method = "t"), nsim = 20)
fix_visit 

```


### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- [GitHub] if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}

#for interaction model
HypesModel
fixef(HypesModel)["DiagnosisTD:VISIT"] <- 0.4 #least effect of interest. 

library(simr)
powerCurveV = powerCurve(HypesModel,
simr::fixed("DiagnosisTD:VISIT", method = "t"), along="SUBJ",nsim = 10)

powerCurveV 
plot(powerCurveV)

#for fixed model 

PypesModel
fixef(PypesModel)["DiagnosisTD"] <- 0.2
fixef(PypesModel)["VISIT"] <- 0.1
PypesModel

powerCurve1 = powerCurve(PypesModel, simr::fixed("DiagnosisTD", method = "t"),along="SUBJ", nsim = 10)

powerCurve1 
plot(powerCurve1)

powerCurve2 = powerCurve(PypesModel, simr::fixed("VISIT", method = "t"), along = "SUBJ", nsim = 10)

powerCurve2
plot(powerCurve2)

#simulating more participants 
#I think we should use makeLmer()

```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why
